---
title: "classifier_GAMs"
output: 
  pdf_document:
    latex_engine: xelatex
header-includes:
  - "\\usepackage{fontspec}"
  - "\\usepackage{xeCJK}"
---


```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

# Import libraries and data

```{r, echo= FALSE}
df_clf = read.csv('/Users/yameiwang/Project/mason_project/EMNLP2023_classifiers/data/clf_noun_structure.csv')
df_mod = read.csv('/Users/yameiwang/Project/mason_project/EMNLP2023_classifiers/data/clf_mod_noun_structure.csv')
print(sprintf("the dimensions of df_clf is %d x %d", dim(df_clf)[1], dim(df_clf)[2]))
print(sprintf("the dimensions of df_mod is %d x %d", dim(df_mod)[1], dim(df_mod)[2]))
```

```{r, warning = FALSE, message=FALSE}
library(mgcv)
library(itsadug)
library(dplyr)
library(emmeans)
library(reticulate)

df_clf$noun1Log = log(df_clf$noun1_freq)
df_clf$noun2Log = log(df_clf$noun2_freq)
df_clf$log_freq_sum <- df_clf$noun1Log + df_clf$noun2Log
df_clf1 <- subset(df_clf, pmi >=0)
# checked sim ==1, which are not true 1
df_clf1 <- subset(df_clf1, sim !=1)
dim(df_clf1) # 8548     13

df_mod$noun1Log = log(df_mod$noun1_freq)
df_mod$noun2Log = log(df_mod$noun2_freq)
df_mod$log_freq_sum <- df_mod$noun1Log + df_mod$noun2Log
df_mod1 <- subset(df_mod, pmi >=0)
dim(df_mod1) # 5778    13

# combine two datasets into a full dataset
combined_df <- rbind(df_clf1, df_mod1)
combined_df$log_freq_sum <- combined_df$noun1Log + combined_df$noun2Log
dim(combined_df)
```

# The models in this section is run by the pretrained fasttext model

## two-way interactions

```{r,warning=FALSE,fig.width = 6, fig.height= 6}
par(mfrow=c(2,2)) 
model1 <- bam(sum_prop ~ te(pmi,sim),data=combined_df,family=gaussian,model='fREML')
pvisgam(model1, view=c("sim","pmi"),hide.label = TRUE,labcex=0.8,main='SCM: full dataset')

# with different pmis
combined_df1 <- rbind(df_clf,df_mod)
combined_df1a <- subset(combined_df1, pmi_win3 >=0 & sim != 1)
combined_df1b <- subset(combined_df1, pmi_win5 >= 0 & sim != 1)
combined_df1c <- subset(combined_df1, pmi_win10 >= 0 & sim != 1)

model1.1 <- bam(sum_prop ~ te(pmi_win3,sim),data=combined_df1a,family=gaussian,model='fREML')
pvisgam(model1.1, view=c("sim","pmi_win3"),hide.label = TRUE,labcex=0.8,main='SCM: full dataset')

model1.2 <- bam(sum_prop ~ te(pmi_win5,sim),data=combined_df1b,family=gaussian,model='fREML')
pvisgam(model1.2, view=c("sim","pmi_win5"),hide.label = TRUE,labcex=0.8,main='SCM: full dataset')

model1.3 <- bam(sum_prop ~ te(pmi_win10,sim),data=combined_df1c,family=gaussian,model='fREML')
pvisgam(model1.3, view=c("sim","pmi_win10"),hide.label = TRUE,labcex=0.8,main='SCM: full dataset')

par(mfrow=c(1,1))

model2 <- bam(sum_prop ~ te(noun1Log,noun2Log),data=combined_df,family=gaussian,model='fREML')
pvisgam(model2, view=c("noun1Log","noun2Log"),hide.label = TRUE,labcex=0.8,main='SCM: full dataset')
```

## three-way interactions

```{r,warning=FALSE,fig.width = 6, fig.height= 6}

model3 = bam(sum_prop~te(pmi,sim,log_freq_sum),
                      data=combined_df,family=gaussian, method='fREML')
plot(model3, main = 'SCM: full dataset')


model4 = bam(sum_prop~te(noun1Log, noun2Log,sim),
                      data=combined_df,family=gaussian, method='fREML')
plot(model4, main = 'SCM: full dataset')

model5 = bam(sum_prop~te(noun1Log, noun2Log,sim),
                      data=df_mod1,family=gaussian, method='fREML')
plot(model5, main = 'SCM: clf_mod_n structure')

model6 = bam(sum_prop~te(noun1Log, noun2Log,sim),
                      data=df_clf1,family=gaussian, method='fREML')
plot(model6, main = 'SCM: clf_n structure')

model7 = bam(sum_prop~te(pmi,sim,log_freq_sum),
                      data=df_clf1,family=gaussian, method='fREML')
plot(model7, main = 'clf_n structure')

model8 = bam(sum_prop~te(pmi,sim,log_freq_sum),
                      data=df_mod1,family=gaussian, method='fREML')
plot(model8, main = 'clf_mod_n structure')
```

# The models in this section is run by the customed fasttext model

## two-way interactions

```{r,warning=FALSE,fig.width = 6, fig.height= 6}
model1a <- bam(sum_prop ~ te(pmi,sim_custom_fasttext),data=combined_df,family=gaussian,model='fREML')
pvisgam(model1a, view=c("sim_custom_fasttext","pmi"),hide.label = TRUE,labcex=0.8,main='SCM: full dataset')
```

## three-way interactions

```{r,warning=FALSE,fig.width = 6, fig.height= 6}

model3a = bam(sum_prop~te(pmi,sim_custom_fasttext,log_freq_sum),
                      data=combined_df,family=gaussian, method='fREML')
plot(model3a, main = 'SCM: full dataset')


model4a = bam(sum_prop~te(noun1Log, noun2Log,sim_custom_fasttext),
                      data=combined_df,family=gaussian, method='fREML')
plot(model4a, main = 'SCM: full dataset')

model5a = bam(sum_prop~te(noun1Log, noun2Log,sim_custom_fasttext),
                      data=df_mod1,family=gaussian, method='fREML')
plot(model5a, main = 'SCM: clf_mod_n structure')

model6a = bam(sum_prop~te(noun1Log, noun2Log,sim_custom_fasttext),
                      data=df_clf1,family=gaussian, method='fREML')
plot(model6a, main = 'SCM: clf_n structure')

model7a = bam(sum_prop~te(pmi,sim_custom_fasttext,log_freq_sum),
                      data=df_clf1,family=gaussian, method='fREML')
plot(model7a, main = 'clf_n structure')

model8a = bam(sum_prop~te(pmi,sim_custom_fasttext,log_freq_sum),
                      data=df_mod1,family=gaussian, method='fREML')
plot(model8a, main = 'clf_mod_n structure')
```

# The models in this section is run by the customed word2vec model

## change the dataframe, and remove unknown nouns
```{r, warning=FALSE}
df_clf2 <- subset(df_clf1, !(noun1 == '堆' | noun2 == '堆'))
df_mod2 <- subset(df_mod1, !(noun1 == '品' | noun2 == '品'))
combined_df2 <- rbind(df_clf2, df_mod2)
```

## two-way interactions

```{r,warning=FALSE,fig.width = 6, fig.height= 6}
model1b <- bam(sum_prop ~ te(pmi,sim_custom_word2vec),data=combined_df2,family=gaussian,model='fREML')
pvisgam(model1b, view=c("sim_custom_word2vec","pmi"),hide.label = TRUE,labcex=0.8,main='SCM: full dataset')
```

## three-way interactions

```{r,warning=FALSE,fig.width = 6, fig.height= 6}

model3b = bam(sum_prop~te(pmi,sim_custom_word2vec,log_freq_sum),
                      data=combined_df2,family=gaussian, method='fREML')
plot(model3b, main = 'SCM: full dataset')


model4b = bam(sum_prop~te(noun1Log, noun2Log,sim_custom_word2vec),
                      data=combined_df2,family=gaussian, method='fREML')
plot(model4b, main = 'SCM: full dataset')

model5b = bam(sum_prop~te(noun1Log, noun2Log,sim_custom_word2vec),
                      data=df_mod2,family=gaussian, method='fREML')
plot(model5b, main = 'SCM: clf_mod_n structure')

model6b = bam(sum_prop~te(noun1Log, noun2Log,sim_custom_word2vec),
                      data=df_clf2,family=gaussian, method='fREML')
plot(model6b, main = 'SCM: clf_n structure')

model7b = bam(sum_prop~te(pmi,sim_custom_word2vec,log_freq_sum),
                      data=df_clf2,family=gaussian, method='fREML')
plot(model7b, main = 'clf_n structure')

model8b = bam(sum_prop~te(pmi,sim_custom_word2vec,log_freq_sum),
                      data=df_mod2,family=gaussian, method='fREML')
plot(model8b, main = 'clf_mod_n structure')
```


